---
title: "Plot Normal Distribution"
author: "Eduardo Capanema"
date: "7 1 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Normal Distribution

In a random collection of data from independent sources, it is generally observed that the distribution of data is normal. Which means, on plotting a graph with the value of the variable in the horizontal axis and the count of the values in the vertical axis we get a bell shape curve. The center of the curve represents the mean of the data set. In the graph, fifty percent of values lie to the left of the mean and the other fifty percent lie to the right of the graph. This is referred as normal distribution in statistics.

R has four in built functions to generate normal distribution. They are described below.

dnorm(x, mean, sd)
pnorm(x, mean, sd)
qnorm(p, mean, sd)
rnorm(n, mean, sd)

Following is the description of the parameters used in above functions −

* x is a vector of numbers.

* p is a vector of probabilities.

* n is number of observations(sample size).

* mean is the mean value of the sample data. It's default value is zero.

* sd is the standard deviation. It's default value is 1.

### dnorm()

This function gives height of the probability distribution at each point for a given mean and standard deviation.

```{r normal}

# Create a sequence of numbers between -10 and 10 incrementing by 0.1.
x <- seq(-10, 10, by = .1)

# Choose the mean as 2.5 and standard deviation as 0.5.
y <- dnorm(x, mean = 2.5, sd = 0.5)

# Give the chart file a name.
plot(x,y)


```


### pnorm()

This function gives the probability of a normally distributed random number to be less that the value of a given number. It is also called "Cumulative Distribution Function".


```{r CDF/FDA}

# Create a sequence of numbers between -10 and 10 incrementing by 0.2.
x <- seq( -10, 10, by = .2 )
 
# Choose the mean as 2.5 and standard deviation as 2. 
y <- pnorm(x, mean = 2.5, sd = 2)


# Plot the graph.
plot(x,y)

```


### qnorm()

This function takes the probability value and gives a number whose cumulative value matches the probability value.

```{r qnorm}
# Create a sequence of probability values incrementing by 0.02.
x <- seq(0, 1, by = 0.02)

# Choose the mean as 2 and standard deviation as 3.
y <- qnorm(x, mean = 2, sd = 1)

# Plot the graph.
plot(x,y)

```


### rnorm()

This function is used to generate random numbers whose distribution is normal. It takes the sample size as input and generates that many random numbers. We draw a histogram to show the distribution of the generated numbers.

```{r rnorm}
# Create a sample of 50 numbers which are normally distributed.
y <- rnorm(50)

# Plot the histogram for this sample.
hist(y, main = "Normal DIstribution")

```


### More Probability Plots

This section describes creating probability plots in R for both didactic purposes and for data analyses.

Probability Plots for Teaching and Demonstration
When I was a college professor teaching statistics, I used to have to draw normal distributions by hand. They always came out looking like bunny rabbits. What can I say?

R makes it easy to draw probability distributions and demonstrate statistical concepts. Some of the more common probability distributions available in R are given below.

distribution	R name	distribution	R name
Beta	beta	Lognormal	lnorm
Binomial	binom	Negative Binomial	nbinom
Cauchy	cauchy	Normal	norm
Chisquare	chisq	Poisson	pois
Exponential	exp	Student t	t
F	f	Uniform	unif
Gamma	gamma	Tukey	tukey
Geometric	geom	Weibull	weib
Hypergeometric	hyper	Wilcoxon	wilcox
Logistic	logis	 	 
For a comprehensive list, see Statistical Distributions on the R wiki. The functions available for each distribution follow this format:

name	description
dname( )	density or probability function
pname( )	cumulative density function
qname( )	quantile function
Rname( )	random deviates
For example, pnorm(0) =0.5 (the area under the standard normal curve to the left of zero). qnorm(0.9) = 1.28 (1.28 is the 90th percentile of the standard normal distribution). rnorm(100) generates 100 random deviates from a standard normal distribution.

Each function has parameters specific to that distribution. For example, rnorm(100, m=50, sd=10) generates 100 random deviates from a normal distribution with mean 50 and standard deviation 10.

You can use these functions to demonstrate various aspects of probability distributions. Two common examples are given below.


```{r plotting}
# Display the Student's t distributions with various
# degrees of freedom and compare to the normal distribution

x <- seq( -4, 4, length=100 )
hx <- dnorm( x )

degf <- c( 1, 3, 8, 30 )
colors <- c( "red", "blue", "darkgreen", "gold", "black" )
labels <- c( "df=1", "df=3", "df=8", "df=30", "normal" )

plot( x, hx, type="l", lty=2, xlab="x value", ylab="Density", main="Comparison of t Distributions" )

for( i in 1:4 ) {
  lines( x, dt( x, degf[i] ), lwd=2, col=colors[i] )
}

legend( "topright", inset=.05, title="Distributions", labels, lwd=2, lty=c( 1, 1, 1, 1, 2 ), col=colors )


# Children's IQ scores are normally distributed with a
# mean of 100 and a standard deviation of 15. What
# proportion of children are expected to have an IQ between
# 80 and 120?

mean=100; sd=15
lb=80; ub=120

x <- seq( -4, 4, length=100 ) * sd + mean
hx <- dnorm( x, mean, sd )

plot( x, hx, type="n", xlab="IQ Values", ylab="", main="Normal Distribution", axes=FALSE )

i <- x >= lb & x <= ub
lines( x, hx )
polygon( c( lb, x[i], ub ), c( 0, hx[i], 0 ), col="red" )

area <- pnorm( ub, mean, sd ) - pnorm( lb, mean, sd )
result <- paste( "P(", lb, "< IQ <", ub, ") =", signif( area, digits=3 ) )
mtext( result, 3 )

axis( 1, at=seq( 40, 160, 20 ), pos=0 )

```



### Fitting Distributions

There are several methods of fitting distributions in R. Here are some options.

You can use the qqnorm( ) function to create a Quantile-Quantile plot evaluating the fit of sample data to the normal distribution. More generally, the qqplot( ) function creates a Quantile-Quantile plot for any theoretical distribution.

```{r fitting}

# Q-Q plots
par( mfrow=c( 1, 2 ) )

# create sample data
x <- rt( 100, df=3 )

# normal fit
qqnorm( x ); qqline( x )

# t(3Df) fit
qqplot( rt( 1000, df=3 ), x, main="t(3) Q-Q Plot", ylab="Sample Quantiles" )
abline( 0, 1 )

```


The fitdistr( ) function in the MASS package provides maximum-likelihood fitting of univariate distributions. The format is fitdistr(x, densityfunction) where x is the sample data and densityfunction is one of the following: "beta", "cauchy", "chi-squared", "exponential", "f", "gamma", "geometric", "log-normal", "lognormal", "logistic", "negative binomial", "normal", "Poisson", "t" or "weibull".

```{r fitdirts}

# Estimate parameters assuming log-Normal distribution

# create some sample data
x <- rlnorm( 100 )

# estimate paramters
library( MASS )
fitdistr( x, "lognormal" )

```





